---
output:
  pdf_document: default
  html_document: default
---
<center><h1>"HEDA" Package</h1></center>

***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<strong>Type</strong>			Package

<strong>Title</strong>    **H**ydropeaking **E**vents **D**etection **A**lgorithm

<strong>Version</strong> 1.0.0

<strong>Author</strong> Tingyu Li, Xiaotian Zou, Gregory B. Pasternack

<strong>Maintainer</strong> Tingyu Li (styli<!-- breaklink -->@ucdavis.edu)

<strong>Description</strong> HEDA is designed to identify hydropeaking events from raw time-series flow record, a rapid flow variantion induced by the hourly-adjusted electricity market. The novelty of HEDA is to use vector angle instead of the first-order derivative to detect change points which not only largely improves the computing efficiency but also accounts for the rate of change of the flow variation. [More details](https://doi.org/10.1016/j.jhydrol.2021.126392)

<strong>License</strong>

<strong>Encoding</strong> UTF-8

<strong>LazyData</strong> true

<strong>RoxygenNote</strong>

<strong>NeedsCompliation</strong> no

<strong>Repository</strong> CRAN

<strong>Date/Publication</strong> 2021-05-21

***

## User Guide
Prior to using this library, four packages needs to install: <strong>dplyr</strong>, <strong>lubridate</strong>, <strong>zoo</strong>, <strong>ggplot2</strong>.
```{r eval=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(zoo)
library(lubridate)
library(HEDA)
```

The input files should be .csv files, and are required to only contain 3 columns, which are site id, datetime and parameter value respectively. Site id is the identification number of gauge stations. Datetime is the date and time of the record. Parameter value refers to the discharge of flow record. 
```{r eval=TRUE}
# import the flow data
HPK_Q <- HPK_Sample_data
head(HPK_Q)
```
***

## Functions
```
HEDA_Tidy        Preprocess the data
```
<strong>Description</strong>

&emsp;Format flow record into hourly record; Split record by designated season; Interpolate and smooth the record.  

<strong>Usage</strong>

&emsp;<span style= "font-family:Verdana">HEDA_Tidy(dataframe, season)</span>

<strong>Arguments</strong>

&emsp; <span style= "font-family:Verdana"> dataframe </span> &emsp;&emsp;&emsp; Name of the dataframe to be processed. </br>
&emsp; <span style= "font-family:Verdana"> season </span> &emsp; Subset data by season.</br>


<strong>Outputs</strong>

&emsp; Output dataframe has 4 columns containing location id, datatime, parameter value, ann_thre. ann_thre is originally the mean annual discharge. If flow record is subset by season, ann_thre will be the mean discharge of that period.

<strong>Example</strong>

``` {r eval=TRUE}
# format the time column to time format
HPK_Q$dateTime <- parse_date_time(HPK_Q$dateTime, "mdy HM")

# clean subset data by season
hpk_flow_cln = HEDA_Tidy(HPK_Q, season = c(6,7,8,9))
head(hpk_flow_cln)
```
***

```
ReversalCount        Detect hydropeaking events
```
<strong>Description</strong>

&emsp; Detect change points of hydropeaking events and classified change points into four categories.

<strong>Usage</strong>

&emsp; <span style= "font-family:Verdana">ReversalCount(dataframe, alpha1, theta)</span>

<strong>Arguments</strong>

&emsp; <span style= "font-family:Verdana">dataframe</span> &emsp; Name of the dataframe to be processed. </br>
&emsp; <span style= "font-family:Verdana">alpha1</span> &emsp;&emsp;&emsp; Default value : 0.03 </br> 
&emsp; <span style= "font-family:Verdana">theta</span> &emsp;&emsp;&emsp;&emsp; Default value : 60 degree for m^3^/s, 85 degree for cfs. gamma default value: 1.1 m^3^/s or 40 cfs.

<strong>Outputs</strong>

&emsp; Output dataframe has six columns containing location id, datatime, parameter value, ann_thre, vt_degree and dgtag. ann_thre is originally the mean annual discharge. If flow record is subset by season, ann_thre will be the mean discharge of that period. vt_degree is the vector angle between two flow vectors. dgtag is the categorized change points.

<strong>Example</strong>

```{r eval=TRUE}
hpk_flow_cg <- ReversalCount(hpk_flow_cln, alpha1 = 0.03, theta = 85, gamma = 40)
head(hpk_flow_cg)
```

***

```
Clean_position        Exclude change points in wrong position
```

<strong>Description</strong>

&emsp; Change points are excluded if they are in the wrong position. For example, both point 3 and the peak pair represent the peaking discharge whose value (position) should be close to the daily maximum discharge. If the peaking discharge is close to the daily minimum discharge, change points will be removed since they are in the wrong positions.

<strong>Usage</strong>

&emsp; <span style= "font-family:Verdana">Clean_position(dataframe, alpha2)</span>

<strong>Arguments</strong>

&emsp; <span style= "font-family:Verdana">dataframe </span>&emsp;  Name of the dataframe to be processed.</br>
&emsp; <span style= "font-family:Verdana">alpha2 </span>&emsp;&emsp;&emsp; Default value : 0.3. </br>

<strong>Outputs</strong>

&emsp; The output is dataframe.

<strong>Example</strong>

```{r eval=TRUE}
hpk_flow_cg <- Clean_position(hpk_flow_cg, alpha2 = 0.3)
```
***


```
Clean_Spt        Clean repeated points
```

<strong>Description</strong>

&emsp; Clean continuous points of the same type

<strong>Usage</strong>

&emsp; <span style= "font-family:Verdana">clean_Spt(dataframe, alpha3, alpha4) </span>

<strong>Arguments</strong>

&emsp; <span style= "font-family:Verdana">dataframe</span> &emsp; Name of the dataframe to be processed. </br>
&emsp; <span style= "font-family:Verdana">alpha3</span> &emsp;&emsp;&emsp; The default value: 0.7</br>
&emsp; <span style= "font-family:Verdana">alpha4</span> &emsp;&emsp;&emsp; The default value: 0.5</br>

<strong>Outputs</strong>

&emsp; The output file is dataframe.


<strong>Example</strong>

```{r eval=TRUE}
hpk_flow_cg <- Clean_Spt(hpk_flow_cg, alpha3 = 0.7, alpha4 = 0.5)
```

***

```
Clean_conectD        Evaluate difference between peaking and off-peaking discharge
```

<strong>Description</strong>

&emsp; Evaluate whether the difference in discharge between peaking and off-peaking points is qualified to be identified as hydropeaking events.

<strong>Usage</strong>

&emsp; <span style= "font-family:Verdana">clean_conectD(dataframe, alpha3, alpha4)</span>

<strong>Arguments</strong>

&emsp; <span style= "font-family:Verdana">dataframe</span> &emsp; Name of the input dataframe to be processed. </br>
&emsp; <span style= "font-family:Verdana">alpha3</span> &emsp;&emsp;&emsp; The default value: 0.7. </br>
&emsp; <span style= "font-family:Verdana">alpha4</span> &emsp;&emsp;&emsp; The default value: 0.5. </br>

<strong>Outputs</strong>

&emsp; The output will be a dataframe in the same form with the input.


<strong>Example</strong>

```{r eval=TRUE}
hpk_flow_cg <- Clean_conectD(hpk_flow_cg, alpha3 = 0.7, alpha4 = 0.5)
```
***

```
HPK_plot        Plot hydrograph
```

<strong>Description</strong>

&emsp; Plot the hydrograph of the processed data with change points marked by different colors.

<strong>Usage</strong>

&emsp; <span style= "font-family:Verdana">HPK_plot(dataframe)</span>

<strong>Arguments</strong>

&emsp; <span style= "font-family:Verdana">dataframe</span> &emsp; Name of the input dataframe to be processed.

<strong>Outputs</strong>

&emsp; The plot will be presented under the "Plots" tab in RStudio. The plot shows flow in m^3^/s.


<strong>Example</strong>

```{r eval=TRUE, fig.width=14}
## subset the data
tt = hpk_flow_cg[13100:13400,]

# plot the hydrograph of the subset data
HPK_plot(tt)
```
***

```
Hpk_Frq_Mgt        Extract frequency and magnitude metrics
```

<strong>Description</strong>

&emsp; Extract Qpeak, offQpeak, pk_no and pkraio from the identified hydropeaking events. 

<strong>Usage</strong>

&emsp; <span style= "font-family:Verdana">HPK_frq_mgt(dataframe)</span>

<strong>Arguments</strong>

&emsp; <span style= "font-family:Verdana">dataframe</span> &emsp; &emsp;&emsp;&emsp;&emsp;Name of the dataframe to be processed. </br>

<strong>Outputs</strong>

&emsp; Output dataframe includes time-series values of four metrics. Qpeak is the hydropeaking discharge, offQpeak is base flow, pk_no is the daily number of rise and fall process, pkratio is the ratio of days with hydropeaking to the total number of days analyzed 


<strong>Example</strong>

```{r eval=TRUE}
HpkFrqMgt <- HPK_frq_mgt(hpk_flow_cg)
head(HpkFrqMgt)
```


```
Hpk_rt_dur        Extract rate of change and duration related metrics
```

<strong>Description</strong>

&emsp; Extract rate of change and duration related metrics: pk_rtn is the retention of peaking process. offpk_rtn is the retention of base flow. D_rampup/D_ramp is the duration of rise/fall process. RB_Indx_up/RB_Index_dw is the flashness of rise/fall process. Ramp_up/Ramp_dw is the rate of change of rise/fall process. Strange_up/Strange_dw is the standardized rise/fall amplitude.

<strong>Usage</strong>

&emsp; <span style= "font-family:Verdana">HPK_frq_mgt(dataframe)</span>

<strong>Arguments</strong>

&emsp; <span style= "font-family:Verdana">dataframe</span> &emsp; &emsp;&emsp;&emsp;&emsp;Name of the dataframe to be processed. </br>

<strong>Outputs</strong>

&emsp; Output dataframe includes time-series values of four metrics. Qpeak is the hydropeaking discharge, offQpeak is base flow, pk_no is the daily number of rise and fall process, pkratio is the ratio of days with hydropeaking to the total number of days analyzed.


<strong>Example</strong>

```{r eval=FALSE}
HpkRtDur <- HPK_rt_dur(hpk_flow_cg)

# to extract the time-series of one metric withouth missing value
pk_rtn <- HpkRtDur[na.omit(HpkRtDur$pk_rtn),c("location_id", "datetime","pk_rtn")]
```
